Optimized Multi-Frequency Spectra for Numerical Radiative Transfer

### 2011-09-07 doc/paper/ms.tex
Added super high-res parameter files for RT06 #1.
Free of scipy

%---------------------------------------- Declarations -----------------------------------------%

%\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}
%\usepackage{emulateapj5}
%\usepackage{apjfonts}
%
%% Define General Packages
%\usepackage{graphicx}
%\usepackage{pslatex}
%\usepackage{amsmath, amsthm, amssymb}
%\usepackage{natbib}
%\usepackage{epsfig}
%
%% Use utf-8 encoding for foreign characters
%\usepackage[utf8]{inputenc}

@@

\title {Optimized Multi-Frequency Spectra for Numerical Radiative Transfer}
\author{Jordan Mirocha, Stephen Skory, Jack Burns}
\affil{Center for Astrophysics and Space Astronomy, University of Colorado, Campus Box 389, Boulder, CO 80309}
\affil{NASA Lunar Science Institute, NASA Ames Research Center, Moffett Field, CA 94043, USA}
\author{John Wise}
\affil{Georgia Institute of Technology}
\author{Steven Furlanetto}
\affil{University of California at Los Angeles}

@@

Though we will find that multi-frequency radiative transfer is necessary in most cases to accurately calculate the ionization and temperature state of the gas, it can also add significant computational expense.  Because the efficiency with which photons ionize and heat a gas depends on their energy, adding more energy groups will alter the amount of ionization and heating taking place in a given computational grid element.  But, these properties of the gas also dictate the computational time-step.  Generally, it is recommended that radiation time-steps are limited by a maximum fractional change in neutral species abundances so that ionization fronts are propagated at the correct speed (\cite{Shapiro2004}).  As a result, the addition of photons that ionize gas more efficiently will lead to a more expensive calculation since the time-step necessary to restrict changes in the neutral fraction will shrink.  Similarly, radiation-hydrodynamic simulations may be slowed by the inclusion of photons that heat gas efficiently, since higher temperature cells have greater sound speeds, and thus require a smaller time-step.  Expense due to additional energy groups is certainly method-dependent, and may even be implementation-dependent.  For example, including high energy photons with long mean free paths will be very costly for a ray-tracing algorithm, but add no expense to a problem solved by flux-limited diffusion.  In addition, some adaptive ray-tracing implementations require each additional energy group in the source spectrum to be traced independently, in which case the time required by the radiative transfer solver scales roughly as \nnu.  Other implementations are now avoiding this problem by tracing all energy groups at once, though keeping them separate can be advantageous since energy groups with small mean free paths will be completely attenuated at small radii (\cite{Wise2011}).

@@

The contents of this paper fall into three general areas.  In \S2 we will quantitatively assess the accuracy with which multi-frequency calculations reproduce the ionization and heating profiles of continuous spectral energy distributions.  \S3 is devoted to introducing a technique for optimally selecting discrete SED templates so as to minimize the errors between discrete and continuous solutions, and \S4 will present the results obtained with this method.  In \S5, we conclude.  Details concerning the code used for this work, as well as a discussion regarding the factors that influence the solution of radiative transfer calculations, whether it be the temporal resolution, spatial resolution, frequency resolution, or the interplay of all three, can be found in the Appendix.

%Reduce guesswork!  Choose frequency resolution in the same way that one chooses spatial/mass resolution.

>>>>

@@

\title {Optimized Multi-Frequency Spectra for Numerical Radiative Transfer}
\author{Jordan Mirocha$^{\dagger}$, Stephen Skory, Jack Burns}
\affil{Center for Astrophysics and Space Astronomy, University of Colorado, Campus Box 389, Boulder, CO 80309}
\affil{NASA Lunar Science Institute, NASA Ames Research Center, Moffett Field, CA 94043, USA}
\email{$^{\dagger}$jordan.mirocha@colorado.edu}
\author{John Wise}
\affil{Georgia Institute of Technology}
\author{Steven Furlanetto}
\affil{University of California at Los Angeles}

@@

Though we will find that multi-frequency radiative transfer is necessary in most cases to accurately calculate the ionization and temperature state of the gas, it can also add significant computational expense.  Because the efficiency with which photons ionize and heat a gas depends on their energy, adding more energy groups will alter the amount of ionization and heating taking place in a given computational grid element.  But, these properties of the gas also dictate the computational time-step.  Generally, it is recommended that radiation time-steps are limited by a maximum fractional change in neutral species abundances so that ionization fronts are propagated at the correct speed (\cite{Shapiro2004}).  As a result, the addition of photons that ionize gas more efficiently will lead to a more expensive calculation since the time-step necessary to restrict changes in the neutral fraction will shrink.  Similarly, radiation-hydrodynamic simulations may be slowed by the inclusion of photons that heat gas efficiently, since higher temperature cells have greater sound speeds, and thus require a smaller time-step.

Expense due to additional energy groups is certainly method-dependent, and may even be implementation-dependent.  For example, including high energy photons with long mean free paths will be very costly for a ray-tracing algorithm, but add no expense to a problem solved by flux-limited diffusion.  In addition, some adaptive ray-tracing implementations require each additional energy group in the source spectrum to be traced independently, in which case the time required by the radiative transfer solver scales roughly as \nnu.  Other implementations are now avoiding this problem by tracing all energy groups at once, though keeping them separate can be advantageous since energy groups with small mean free paths will be completely attenuated at small radii (\cite{Wise2011}).

@@

The contents of this paper fall into three general areas.  In \S2 we will quantitatively assess the accuracy with which multi-frequency calculations reproduce the ionization and heating profiles of continuous spectral energy distributions.  \S3 is devoted to introducing a technique for optimally selecting discrete SED templates so as to minimize the errors between discrete and continuous solutions, and \S4 will present the results obtained with this method.  In \S5, we conclude.  Verification of the code used for this work can be found in the Appendix.
